\section*{Reviewer 4 comments}
\textit{Comments to the Author}\\

This work presents MERF; a GUI-enabled regular expressions 
based tool for entity and relational entity extraction, 
enhanced for Arabic text by enabling morphology-based matching.

\begin{enumerate}[leftmargin=0mm,label=\bfseries CommentR2.\arabic*]


\item \label{Review.4.1} 
Clarity:   
It is not clear how to associate code actions with 
subexpressions!

\answer{
  We elaborate the use of code actions in Subsection~\ref{subsec:numnorm} 
    by introducing Figure~\ref{fig:numnormalgo} and discussing it 
    in the corresponding case study. 
}

\item \label{Review.4.2} 
Correctness :  
- $Syn^k$ introduces a lot of noise especially if $k$ is large. 

\answer{
  Your note is accurate, this is why we limit the degree (k) to 7.
  We further elaborate on this by introducing the following in the paper; Subsection~\ref{subsec:grammar}.
  \begin{quote}
  We limit $k$ to a maximum of $7$ since we practically noticed that 
  (1) values above $7$ introduce significant semantic noise and
  (2) the computation is expensive without a bound. 
  \end{quote}
}

\item \label{Review.4.3} 
In the results section, there is no standard evaluation. 
The evaluation is based on comparisons with non-standard 
task-specific tools. 

\answer{
  We modified the Results Section~\ref{sec:results} to 
    clarify our evaluation method as follows.
    \begin{itemize}
      \item We compare two different approaches trying to achieve the 
        same results using the same input corpora and the same
        golden reference. 
      \item We use standard metrics to report the results: development
        time, size of solution, runtime, precision and recall.
      \item We use the reported task-specific implementations because, 
        up to our knowledge, no other accessible relational entity 
        extraction tools exist for Arabic. 
    \end{itemize} 
}

\item \label{Review.4.4} 
Also, the reported precision and recall are very sensitive 
to the quality of the annotator!

\answer{
  This is correct. We added a precision discussion paragraph
    to reflect that in the Discussion Section~\ref{sec:discuss}. 
}

\end{enumerate}

