This paper presents Sarf, an application customizable Arabic morphological analyzer. 
NLP applications can implement the Sarf API to 
to refine the morphological analysis on the fly by 
(1) selecting the interesting features, (2) prioritizing the features, and
(3) accepting and rejecting solutions on the fly based on partial features reported 
so far. 
Sarf extends the SAMA and BAMA lexicons and 
represents affixes using agglutinative and fusional morphemes which 
(1) significantly reduces the size of the lexicons needed to represent Arabic morphology,(2) fixes inconsistencies in morphological features corresponding to morphemes,
(3) simplifies the maintenance and augmentation of the affix morpheme lexicons, and
(4) solves the segmentation correspondence problem between the morphological solution 
and the original text. 
Sarf also allows the NLP application to use partial diacritics for morphological solution
disambiguation, and solves the `run-on words' problem.
Sarf is available online as an open source tool. 
\todo{R2.17}
We evaluated Sarf in terms of effectiveness and performance with respect to 
other tools such as SAMA, BAMA and ElixirFM. Sarf outperforms other analyzers in 
run-time and slightly increases recall rates in morphological analysis due to the 
extended lexicons. 
We used Sarf application specific API in several NLP applications 
for information extraction from Arabic documents and the applications
provided more accurate results than existing analyzers with faster running time. 

In the future, we plan to improve Sarf by allowing root analysis of stems,
supporting inflectional stems, and providing a graphical user interface to 
allow the users to edit the affix and stem morpheme lexicons of Sarf. 
