\section*{Editor Comments to the Author}
Dear Prof. Zaraket,

We have now received all reviewers’ reports on your paper ID NLE-ARTC-15-0011 entitled ``Sarf: Efficient and Application Customizable Arabic Morphological Analyzer".

As you can see from the reviews enclosed, the reviewers identified a number of issues in your article.
As a consequence, we cannot accept your paper as it stands; however, WE ENCOURAGE YOU TO REVISE YOUR ARTICLE according to reviewers' feedback and RESUBMIT IT. (The comments of the reviewers are included at the bottom of this letter).

If you decide to resubmit, please enclose a list of the points that have been addressed, and the criticisms which they answer.

Once we have received your resubmission, it will be checked and will also undergo additional reviews which may require further modification. This is to ensure the paper meets the standards of the journal.

I would be grateful if you could let me know for administrative purposes whether you are prepared to make the suggested amendments, and, if so, when I can expect to receive the revised version.

To revise your manuscript, log into \url{https://mc.manuscriptcentral.com/nle} and enter your Author Center, where you will find your manuscript under ``Manuscripts with Decisions."  Under "Actions," click on "Create a Revision."  Your manuscript number will be appended to denote a revision. You may also click this link to start your revision:

\url{https://mc.manuscriptcentral.com/nle?URL_MASK=f5a4558b88a24181a82b5db451930250}

When submitting your revised manuscript, you will be able to respond to the comments made by the reviewers in the space provided.  Please use this space to document any changes you make to the original manuscript.

Could you please acknowledge receipt of this e-mail?

We hope that the reviewers' feedback will be useful for you in improving your work.

Thank you again for submitting your article to the NLE.

Best regards,

