{\em Computational Linguistics} (CL) is concerned with building 
accurate linguistic computational models.
{\em Natural Language Processing} (NLP) is concerned with automating the 
understanding of natural language. 
CL and NLP tasks range from simple ones such as spell checking and typing error correction to more complex tasks including %text summarization, 
{\em named entity recognition} (NER), {\em cross-document analysis}, 
{\em machine translation}, and {\em relational entity extraction}~\cite{linckels2011natural,ferilli2011natural}.
Entities are elements of text that are of interest to an NLP task.
Relational entities are elements that connect entities.
{\em Annotations} relate chunks of text to {\em labels} denoting 
semantic values such as entities or relational entities.
We refer to annotations and labels as {\em tags} and {\em tag types}, 
respectively, in the sequel.

Supervised and unsupervised empirical learning techniques tackle NLP and CL tasks.
They employ machine learning without the need to manually encode the requisite knowledge~\cite{soudi2007arabic}. 
Supervised learning techniques require training corpora annotated 
with {\em correct} tags to learn a computational model.
Supervised and unsupervised techniques require annotated reference corpora to 
evaluate the accuracy of the technique using metrics such as precision and 
recall~\cite{englishtreebank,arabictreebank,chinesetreebank}. 

Researchers build %the required %seed 
training and reference corpora 
either manually, incrementally using learning techniques, or using knowledge-based 
annotation techniques that recognize and extract entities and relational 
entities from text. 
Knowledge-based techniques use linguistic and 
rhetorical domain specific knowledge encoded into sets of rules 
to extract entities and relational entities~\cite{soudi2007arabic}.
While existing annotation, entity, and relational entity 
extraction tools exist
\cite{chiticariu2010systemt% SystemT
,atzmueller2008rule%Textmarker
,urbain2012user%
,settles2011closing% dualist
,mmax2%MMAX
,brat% BRAT
},
most of them lack Arabic language support, and almost all of them
lack Arabic morphological analysis support
% that is key to Arabic NLP due to the rich morphological nature of Arabic
~\cite{habash2006arabic}. 
Fassieh~\cite{attia2009fassieh} is a {\em commercial} Arabic annotation tool
with morphological analysis support and text factorization. 
However, this tool lacks support for entity and relational entity extraction.

\input{directionFigure}

Figure~\ref{fig:intromotiv} illustrates the target of \framework using 
the directions to Dubai Mall example~\footnote{text taken from the Dubai 
Mall website ~\url{http://www.thedubaimall.com/ar/}}.
The figure also presents a transliteration and an English translation of 
the Arabic text.
The framed words in the text are entities referring to names of people 
($n_1,n_2,n_3$), 
names of places ($p_1,\dots,p_7$), 
relative positions ($r_1,\dots,r_4$), 
and numerical terms ($u_1,u_2$). 
We would like to extract those entities, and then extract
the relational entities forming the graph in Figure~\ref{fig:intromotiv} 
where vertices express entities, 
and edges represent the relational entities.


%These tasks take as input digital text documents from sources including literature, books, news, business reports, chat messages, and emails. 
%Some documents are originally typed in digital format such as emails. 
%Other documents are automatically digitized using techniques such as {\em optical character recognition} (OCR) and 
%{\em automatic speech recognition}~\cite{margner2012guide,rashwan2012robust}. 

In this paper, 
we present \framework, a morphology-based entity and relational entity
extraction framework for Arabic text.
\framework provides a user-friendly interface where the user defines tag types 
and associates them with
%\framework formulae that are 
regular expressions over 
%\framework
Boolean formulae.
%
A Boolean formula is defined by a term, negation of a term, or disjunction of terms.
Terms are matches to Arabic morphological features including 
prefix, stem, suffix, part of speech (POS) tags, gloss tags, extended synonym 
tags, and semantic categories.
For example, entity $p_1$ in Figure~\ref{fig:intromotiv} has a place semantic 
category.
\framework regular expressions support operators such as concatenation, 
zero or one, zero or more, one or more, up to $M$ repetitions where $M$ is 
a non-zero positive integer, and logical conjunction and disjunction.
For example, the sequence between $p_1$ and $p_2$ matches a regular expression $re$ 
that requires two semantic place categories with a place-preposition POS 
tag ($r_1$) in between.

An editor allows the user to associate an action with each sub-expression.
The user specifies the action with C++ code and uses an API to access information 
related to the matches such as text, position, length, morphological features, 
and numerical value.
%
Each regular expression is associated with a named identifier to form a 
{\em local grammar} like structure~\cite{traboulsi2009arabic}. 
%
A relation definition GUI allows the user to provide relational tuples 
where each tuple has a 
source, a destination and an edge label.
The user uses the regular expression identifiers to define the relational 
tuple elements.
For example, the relation between $e_1, e_2$ and $r$ shown in 
Figure~\ref{fig:intromotiv} is a match of a relational tuple 
over the components of $re$. 
%
We refer to regular expressions and Boolean formulae as expressions and formulae, respectively.
We also refer to expressions as rules
when used in a grammar context; e.g.  when used with an identifier. 
%We discuss the importance of the morphological features supported in 
%\framework terms in Section~\ref{sec:morph}; in brief, morphological preprocessing
%is key to Arabic NLP.% due to the richness of Arabic morphology.

\framework takes an Arabic text and the local grammar defined by the 
Boolean formulae and the regular expressions.
\framework computes the morphological solutions of the 
input text then computes matches to the Boolean formulae therein.
\framework then generates a {\em non-deterministic finite state 
automata} (NDFSA) for each expression and simulates it with the 
sequence of Boolean formulae matches to compute the regular 
expression matches. 
\framework generates executable code for the actions associated with
the regular expressions, 
compiles, links, and executes the generated code 
as shared object libraries.
Finally, \framework constructs the semantic relations and 
cross-reference between entities.
\framework also provides visualization tools to present the matches, 
and estimate their accuracy with respect to reference tags.

This work significantly extends~\cite{JaZaMatar} 
that allows for manual, and morphology annotation.
\framework enables a user to incrementally create 
complex annotations for Arabic based on automatic 
extraction of morphological tags through a user friendly interactive interface. 
\framework has the following advantages.
\begin{itemize}
  \item \framework provides a novel and intuitive visual interface to build formulae over morphological features, 
    build regular expressions over the resulting formulae, and thereafter compute automatic tags.
  \item To our knowledge, this morphology-based framework is the first for Arabic entity and relational entity extraction.
  \item \framework provides the user with the ability to rapidly create annotated Arabic text corpora with sophisticated morphology-based tags.
\end{itemize}

In \framework, we make the following contributions.
\begin{itemize}
  \item \framework enables the user to define relations in a simple manner
    and automatically detects relational entities matching the user defined relations. 
  \item \framework enables the user to associate sub-expressions
    with code actions, and executes the code action 
    when a corresponding match is found.
    It also provides an API to enable access to match 
    features such as text, position, length, numerical value, and morphological features.
  \item \framework enables the user to tag words based on a novel light Arabic WordNet relation that leverages the synonym $Syn^k$ feature.
  \item \framework is open source and available online for the research community 
    under on (\url{https://github.com/codelogicanalysis/atmine}).
\end{itemize}

The rest of the paper is structured as follows.
%In the rest of this paper,
%Section~\ref{sec:motivation} presents a motivational example. 
Section~\ref{sec:morph} introduces Arabic morphological analysis 
and its important role in Arabic NLP.
%We present preliminaries necessary to understand the work 
%in Section~\ref{sec:preliminaries}. 
Section~\ref{sec:methodology} explains the methodology of \framework.
Section~\ref{sec:framework} presents \framework components.
Section~\ref{sec:gui} presents \framework GUI. 
Section~\ref{sec:related} presents and discusses related work.
Section~\ref{sec:results} presents the evaluation results. 
Finally, we conclude and discuss future work in Section~\ref{sec:conclusion}.